Aa Language Reference Manual

### 2013-02-19 v2/Ahir-asm/devel/docs/AaLRM.tex
further progress.. the end is neeeaaaarrrrr...
Linearity checker example completed.  Also started updating docs (long overdue).

\end{verbatim}
For example:
\begin{verbatim}
$if (a != 0) $then
   q := (r + s)
   t := 0
$else
   qdash :=  (r - s)
$endif
\end{verbatim}
If the control token reaches the end of a selected
segment in the if statement (that is, without being
rerouted by a place statement), then the control
token is passed on to the statement immediately
following the if statement.
\item The place statement: The {\bf place} statement
identifies a place which will contain 
a token after the place statemet has executed.  For example,
\begin{verbatim}
$place [fastpath]
\end{verbatim}%$
means that the incoming token is placed in a labeled
place {\bf fastpath} (the place statement never
puts a token into its default exit place).  
The token placed in {\bf fastpath} must be used to
trigger a unique {\bf merge} statement which is required
to depend on this labeled place in the same branch
block. If no such merge exists, or if multiple such merges exist, 
then this is an error.
\item Merge statements: A merge statement is
specified as 
\begin{verbatim}
merge-statement :=
          merge-spec label-list merge-assignments end-merge-spec
\end{verbatim}
which is to be interpreted as follows.  The labels in the label list
refer to token labels defined by {\bf place}  statements within the branch block.
Whenever a token is present in any of the places in
the label-list for a merge,
the merge statement starts and executes a series of special assignments
which multiplex values into targets based on which arc the
token arrived from.  The merge statement then releases its token to
the next statement. 
\begin{itemize}
\item 
The merge assignments inside a merge block are all of a specific
form, called the phi-statement.  A phi-statement has the following
form
\begin{verbatim}
phi-spec target-expr := expr1 on-spec place-spec-1  [expr2 on-spec place-spec-2 ... ]
\end{verbatim}
For example:
\begin{verbatim}
$phi a := b $on $entry c $on loopback 
\end{verbatim}
This says that the target $a$ is to take the value of $b$ if
the merge statement was started from its entry place (from a
token passed from its predecessor) and is to take the value of $c$
if the merge statement was started from the place ``loopback''.
Note that if the merge statement is to be triggered by a
token in its default entry place, one of the labels in the
label-list for the merge must contain the identifier \$entry.
\end{itemize}
\end{itemize}

@@

>>>>

\end{verbatim}
For example:
\begin{verbatim}
$if (a != 0) $then
   q := (r + s)
   t := 0
$else
   qdash :=  (r - s)
$endif
\end{verbatim}
If the control token reaches the end of a selected
segment in the if statement (that is, without being
rerouted by a place statement), then the control
token is passed on to the statement immediately
following the if statement.
\item The place statement: The {\bf place} statement
identifies a place which will contain 
a token after the place statemet has executed.  For example,
\begin{verbatim}
$place [fastpath]
\end{verbatim}%$
means that the incoming token is placed in a labeled
place {\bf fastpath} (the place statement never
puts a token into its default exit place).  
The token placed in {\bf fastpath} must be used to
trigger a unique {\bf merge} statement which is required
to depend on this labeled place in the same branch
block. If no such merge exists, or if multiple such merges exist, 
then this is an error.
\item Merge statements: A merge statement is
specified as 
\begin{verbatim}
merge-statement :=
          merge-spec label-list merge-assignments end-merge-spec
\end{verbatim}
which is to be interpreted as follows.  The labels in the label list
refer to token labels defined by {\bf place}  statements within the branch block.
Whenever a token is present in any of the places in
the label-list for a merge,
the merge statement starts and executes a series of special assignments
which multiplex values into targets based on which arc the
token arrived from.  The merge statement then releases its token to
the next statement. 
\begin{itemize}
\item 
The merge assignments inside a merge block are all of a specific
form, called the phi-statement.  A phi-statement has the following
form
\begin{verbatim}
phi-spec target-expr := expr1 on-spec place-spec-1  [expr2 on-spec place-spec-2 ... ]
\end{verbatim}
For example:
\begin{verbatim}
$phi a := b $on $entry c $on loopback 
\end{verbatim}
This says that the target $a$ is to take the value of $b$ if
the merge statement was started from its entry place (from a
token passed from its predecessor) and is to take the value of $c$
if the merge statement was started from the place ``loopback''.
Note that if the merge statement is to be triggered by a
token in its default entry place, one of the labels in the
label-list for the merge must contain the identifier \$entry.
The target of a phi-statement must be an implicit variable
(declared by the statement itself, not a storage or a pipe
object).
\end{itemize}
\end{itemize}

@@

\subsection{Pipelinable-loops: the {\bf do-while} construct}

The \Aa language provides a single construct for describing loops.
The form of the construct is
\begin{verbatim}
  $do
     $merge $entry $loopback
       <phi-statements>
     $endmerge
     atomic-statement-sequence
  $while <condition-expression>
\end{verbatim}
The do-while statement can appear only in branch-blocks.  Control enters
the do-while from a predecessor in the branch-block, and stays in the
do-while as long as the condition-expression is false.  The \$loopback
place is used to repeat the loop-body.  The leading merge statement
in the do-while is used to merge the control-flow coming in either from
the entry to the do-while or from the loopback place.

The downstream tools will always attempt to pipeline the do-while loop by
starting the next iteration while the current one is still in progress.
Dependency analysis on values in the loop-body is used to determine the
extent of loop pipelining.  Note that it is always possible to describe
a do-while loop using a branch-block, and this is what should be done
if loop-pipelining is not desired.
